---
title: "Anomaly  detection"
author: "kelvin njunge"
date: "9/10/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  **Defining the Question**

## a) **Specifying the Question**
The objective of this project fraud detection by checking whether there are any anomalies in the given sales dataset that could point out potential fraud activity.
## b) Defining the Metric for Success
Exhaustively performing anomaly detection without any errors.

## c) **Understanding the context**
Working as a consultant Data analyst at Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax).This project endeavors to  explore a recent marketing dataset and check whether there are any anomalies in the given sales dataset that could point out potential fraud activity.

## d) **Recording the experimental design**
Importing and reading the data
Data Cleaning
Anomalies detection
Conclusions and recommendations

## e) **Data Relevance**
The data was provided by the company (http://bit.ly/CarreFourSalesDataset). 

```{r}
library(future)
library(fracdiff)
library(lmtest)
library(tseries)
library(tidyverse)
library(anomalize)
library(tibbletime)
```


```{r}
getwd()
```

```{r}
df <- read.csv("C:\\Users\\Ricky\\Documents\\Supermarket_Sales_Forecasting - Sales.csv")
```

```{r}
# preview the dataset
head(df)
```

```{r}
#stucture of the dataset
str(df)
```

```{r}
# reformatting the dates and sortings
df$Date <- as.Date(df$Date, format = "%m/%d/%Y")
df$Date <- sort(df$Date,  decreasing = FALSE)

```


```{r}
# casting as a tibble
data <- as_tbl_time(df, index = Date)
# getting unique daily entries without multiple entries
data <- data %>%
  as_period(period = "daily")
# dimensions of data
dim(data)
```

```{r}
# getting and plotting data for anomaly detection
data %>%
  time_decompose(Sales) %>%
  anomalize(remainder) %>%
  time_recompose() %>%
  plot_anomalies(time_recomposed = TRUE, ncol = 3, alpha_dots = 0.5)
```

## Conclusion

There were no anomalies detected in the data.


